\documentclass{hippoidC}

\memoto{Idris}
\memosubject{Book of Proof}
\memodate{2024.03.21}
\status{\S 11.4 Equivalence Classes and Partitions}

\begin{document}
\toc
\thispagestyle{styleTOC}
\pagebreak
\pagestyle{styleE}

\begin{prooflist}{1. List all the partitions of the set $A=\{a, b\}$. Compare your answer to the answer to Exercise 5 of Section 11.3.}
	\item
	\begin{align*}
		\set{\set{a}, \set{b}} \\
		\set{\set{a, b}}
	\end{align*}
\end{prooflist}

\begin{prooflist}{2. List all the partitions of the set $A=\{a, b, c\}$. Compare
		your answer to the answer to Exercise 6 of Section 11.3.}
	\item
	\begin{align*}
		\set{\set{a}, \set{b}, \set{c}} \\
		\set{\set{a, b}, \set{c}}       \\
		\set{\set{a, c}, \set{b}}       \\
		\set{\set{b, c}, \set{a}}       \\
		\set{\set{a, b, c}}
	\end{align*}
\end{prooflist}

\begin{prooflist}{3. Describe the partition of $\mathbb{Z}$ resulting from the equivalence relation $\equiv(\bmod 4)$.}
	\item Let $R$ be the relation $\equiv\bmod 4$.
	\item For any integer $a$, the value of $a\bmod 4$ is in the set $\set{0, 1, 2, 3}$.
	\item Therefore each element in $\set{0, 1, 2, 3}$ defines an equivalence class
	on $\mathbb{Z}$ from the equivalence relation $\equiv(\bmod 4)$.
	\item Thus the 4 equivalence classes are
	\begin{align*}
		\set{-15, -11, -7, -3, 1 \dots x\bmod4=3} \\
		\set{-14, -10, -6, -2, 2 \dots x\bmod4=2} \\
		\set{-13, -9, -4, -1, 3 \dots x\bmod4=1}  \\
		\set{-12, -8, -4, 0, 4 \dots x\bmod4=0}   \\
	\end{align*}

\end{prooflist}

\begin{prooflist}{4. Suppose $P$ is a partition of a set $A$. Define a relation $R$ on $A$ by declaring $x R y$ if and only if $x, y \in X$ for some $X \in P$. Prove $R$ is an equivalence relation on $A$. Then prove that $P$ is the set of equivalence classes of $R$.}
	\item Let $P$ be a partition of set $A$.
	\item Define relation $R$ on $A$ as follows: $x R y$ if and only if $x$ and $y$ belong to the same set $X \in P$.
	\item To prove $R$ is an equivalence relation:
	\begin{itemize}
		\item Reflexivity: For any $x \in A$, since $x$ belongs to some set $X \in P$, $x R x$ by definition. Hence, $R$ is reflexive.
		\item Symmetry: If $x R y$, then $x$ and $y$ belong to the same set $X \in P$. But this also means $y$ and $x$ belong to the same set $X$, so $y R x$. Thus, $R$ is symmetric.
		\item Transitivity: If $x R y$ and $y R z$, then both $x$ and $y$, and $y$ and $z$, belong to the same sets $X$ and $Y$ in $P$ respectively. Since $P$ is a partition, $X$ and $Y$ are either identical or disjoint. If they are identical, then $x R z$ trivially. If they are disjoint, then $x$ and $z$ cannot belong to the same set in $P$, so $x R z$ holds vacuously. In either case, $R$ is transitive.
	\end{itemize}
	\item Now, to prove that $P$ is the set of equivalence classes of $R$:
	\begin{itemize}
		\item Each set $X \in P$ is an equivalence class of $R$: By definition, elements in the same set $X$ under $P$ are related by $R$. So, $P$ consists of equivalence classes of $R$.
		\item Every equivalence class of $R$ is a set in $P$: Since $P$ is a partition, each element of $A$ belongs to exactly one set $X \in P$. Thus, each equivalence class of $R$ corresponds to a set in $P$.
	\end{itemize}
\end{prooflist}

\begin{prooflist}{5. Consider the partition $P=\{\{\ldots,-4,-2,0,2,4,
			\ldots\},\{\ldots,-5,-3,-1,1,3,5, \ldots\}\}$ of $\mathbb{Z}$. Let $R$ be
		the equivalence relation whose equivalence classes are the two elements of $P$.
		What familiar equivalence relation is $R$ ?}
	\item $xRy$ is true when $x$ and $y$ have the same parity.

\end{prooflist}

\begin{prooflist}{6. Consider the partition
		$P=\{\{0\},\{-1,1\},\{-2,2\},\{-3,3\},\{-4,4\}, \ldots\}$ of $\mathbb{Z}$.
		Describe the equivalence relation whose equivalence classes are the elements of
		$P$.}
	\item $xRy$ is defined by $|x|=|y|$.
\end{prooflist}

\end{document}
